---
title: "Regression"
author: "Ran Dou, Mduduzi Langwenya, Kimo Li, Siyan Lin, Muhammad Furqan Shaikh, Tianyi Zhou"
date: "03/09/2019"
output: html_document
---
### Load the packages
```{r}
rm(list=ls())
library(tidyverse)
library(readr)
library(tidyverse)
library(forecast)
library(leaps)
library(pROC)
library(ggplot2)
library(reshape)
library(car)
library(leaps)
library(corrplot)
library(knitr)
library(broom)
```

### I. Data cleaning and impution

##### Data importing
```{r}
###import the raw diabetes data
diabetes <- read_csv("diabetes.csv")
###delete all the missing valuse
diabetes1 <- diabetes %>%
  filter( Glucose !=0 & BMI != 0 & BloodPressure != 0 & Insulin != 0 & SkinThickness != 0) %>%
  select(Glucose, Insulin, Outcome, BMI, SkinThickness )
```

##### Fill-in Zero Value
###### 1) Insulin
```{r}
### Insulin 
# stepwise for choosing models for Insulin 
insu.lm.null <- lm(Insulin~1, data = diabetes1)
insu.lm <- lm(Insulin~., data = diabetes1)
summary(insu.lm.null)
summary(insu.lm)
insu.lm.step_both <- step(insu.lm, direction = "both")
sum_both <- summary(insu.lm.step_both)
### create the model for imputing Insulin missing values
lm.data <- lm (Insulin ~ Glucose + BMI, data=diabetes1)
pred.1 <- predict (lm.data, diabetes1)
impute <-function(a, a.impute){
         ifelse(a$Insulin == 0, round(a.impute, 0), a$Insulin)
}
diabetes$newInsu <- impute(diabetes, pred.1)
```

###### 2) Skinthickness 
```{r}
### stepwise for choosing models for Insulin 
skin.lm.null <- lm(SkinThickness~1, data = diabetes1)
skin.lm <- lm(SkinThickness~., data = diabetes1)
skin.lm.step_both <- step(skin.lm, direction = "both")
sum_both_skin <- summary(skin.lm.step_both)
### create the model for imputing SkinThickness missing values
lm2.data <- lm(SkinThickness ~ BMI, data=diabetes1)
pred.2 <- predict (lm2.data, diabetes1)
impute <-function(a, a.impute){
  ifelse(a$SkinThickness == 0, round(a.impute, 0), a$SkinThickness)
}
diabetes$newSkin <- impute(diabetes, pred.2)

```

***

### II. Descriptive statistics and visualizations

```{r}
### Create theme for plots
theme <- theme_test(base_family = "Times New Roman") + theme(plot.title = element_text(hjust = 0.5), 
         legend.position = "bottom", panel.grid.minor = element_blank(), axis.ticks.x = element_blank(),
         axis.ticks.y = element_blank(), panel.grid.major = element_blank())
```

```{r, fig.width=3}
##### Raw Data
### Descriptive Statistics
melted_diabetes <- diabetes %>%
  gather(Variables, Value, -c(Pregnancies, DiabetesPedigreeFunction, Outcome, newInsu, newSkin))
ggplot(melted_diabetes,aes(x=Value)) +
  geom_histogram()+ facet_wrap(~Variables, scales = "free_x") +
  theme + labs(title = "Graph 1. Features Overview", x = "Features", y = "")
# Delete missing value for other variables
diabetes <- diabetes %>%
  filter(BloodPressure !=0 & BMI !=0, Glucose != 0 & newInsu != 0)
```

```{r}
##### New Data
### Descriptive Statistics
melted_diabetes <- diabetes %>%
  gather(Variables, Value, -c(Pregnancies,DiabetesPedigreeFunction, Outcome, Insulin, Age))
ggplot(melted_diabetes,aes(x=Value)) +
  geom_histogram()+ facet_wrap(~Variables, scales = "free_x") +
  theme + labs(title = "Graph 1. Features Overview", x = "Features", y = "")
# Delete missing value for other variables
diabetes <- diabetes %>%
  filter(BloodPressure !=0 & BMI !=0, Glucose != 0 & newInsu != 0)
```

```{r, fig.width=3}
### Histograms
# Outcome
summary(diabetes)
ggplot(diabetes, aes( x = factor(Outcome))) + 
  geom_histogram(stat = "count", width = 0.3) +
  theme + labs(title = "Graph 2. Distribution of Outcome", x = "Outcome", y = "Count")
```

```{r}
percentage <- diabetes %>% group_by(Outcome) %>% summarize(n = n()) %>% mutate(percentage = n/sum(n))
pie <-ggplot(percentage, aes(x="", y=percentage,fill = factor(Outcome)))
pie=pie+geom_bar(stat="identity", alpha = 0.6)+
  coord_polar(theta="y")+
  geom_text(aes(label = paste0(round(percentage*100, 2), "%")), position = position_stack(vjust = 0.5)) +
  theme(legend.title = element_blank(), axis.text.x=element_blank(), axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.line = element_blank(), panel.background = element_blank()) + 
  labs(x = "", y = "") + 
  scale_fill_manual(values = c("black","red"))
pie
```

##### Create New Features
```{r}
# Create Age Ranges
diabetes_v <- diabetes %>%
  mutate( newAge = ifelse(Age <= 15, "0-15",
                   ifelse(Age > 15 & Age <= 30, "16-30",
                   ifelse(Age > 30 & Age <= 45, "31-45",
                   ifelse(Age > 45 & Age <= 60, "46-60","60+")))))
# Create BMI Ranges
diabetes_v <- diabetes_v %>%
  mutate( newBMI = ifelse(BMI <= 18.5, "Underweight",
                   ifelse(BMI > 18.5 & BMI <= 25, "Normal",
                   ifelse(BMI > 25 & BMI <= 30, "Over Weight",
                          "Obese"))))
diabetes_v$newBMI=factor(diabetes_v$newBMI, 
                       levels=c("Underweight","Normal","Over Weight", "Obese"))
```

##### Visualizations

```{r}
diabetes <- diabetes %>%
  select( Pregnancies, Glucose, BloodPressure,newInsu, newSkin, BMI, DiabetesPedigreeFunction, Age, Outcome)
```

```{r, fig.width=3}
summary(diabetes_v)
sd(diabetes_v$newInsu)
# Target variable = Outcome
ggplot(diabetes_v, aes( x = newInsu)) + 
  geom_histogram() + 
  theme + labs(title = "Graph 3. Distribution of Insulin", x = "Insulin", y = "Count")
```

```{r, fig.width=2}
mean(diabetes_v$Glucose[diabetes_v$Outcome=="0"])
mean(diabetes_v$Glucose[diabetes_v$Outcome=="1"])
mean(diabetes_v$Age[diabetes_v$Outcome=="0"])
mean(diabetes_v$Age[diabetes_v$Outcome=="1"])
mean(diabetes_v$Glucose[diabetes_v$newAge=="16-30"])
mean(diabetes_v$Glucose[diabetes_v$newAge=="31-45"])
mean(diabetes_v$Glucose[diabetes_v$newAge=="46-60"])
mean(diabetes_v$Glucose[diabetes_v$newAge=="60+"])
```

```{r, fig.width=2}
# boxplots glucose by outcome
ggplot(diabetes_v, aes(x = factor(Outcome), y = Glucose)) +
  geom_boxplot(colour = "black", outline = FALSE, width = 0.6) +
  theme + labs(title = "Graph 4. Distribution of Glucose by Outcome", x = "Outcome", y = "Glucose")

# boxplots bloodpressure by outcome
ggplot(diabetes_v, aes(x = factor(Outcome), y = BloodPressure)) +
  geom_boxplot(colour = "black", outline = FALSE) +
  theme + labs(title = "Graph 5. Distribution of BloodPressure by Outcome", x = "Outcome", y = "BloodPressure")

# boxplots Age by outcome 
ggplot(diabetes_v, aes(x = factor(Outcome), y = Age)) +
  geom_boxplot(colour = "black", outline = FALSE) +
  theme + labs(title = "Graph 6. Distribution of Age by Outcome", x = "Outcome", y = "Age")

# boxplots bmi by outcome
ggplot(diabetes_v, aes(x = factor(Outcome), y = BMI)) +
  geom_boxplot(colour = "black", outline = FALSE) +
  theme + labs(title = "Graph 7. Distribution of BMI by Outcome", x = "Outcome", y = "BMI")

# boxplots pedigree by outcome
ggplot(diabetes_v, aes(x = factor(Outcome), y = DiabetesPedigreeFunction)) +
  geom_boxplot(colour = "black", outline = FALSE) +
  theme + labs(title = "Graph 8. Distribution of DPF by Outcome", x = "Outcome", y = "DPF")

# boxplots pedigree by outcome
ggplot(diabetes_v, aes(x = factor(Outcome), y = newInsu)) +
  geom_boxplot(colour = "black", outline = FALSE) +
  theme + labs(title = "Graph 9. Distribution of Insulin by Outcome", x = "Outcome", y = "Insulin")
```


```{r, fig.width=2}
# boxplots Glucose by Age
ggplot(diabetes_v, aes(x = newAge, y = Glucose)) +
  geom_boxplot(colour = "black") +
  theme + labs(title = "Graph 10. Distribution of Glucose by Age", x = "Age", y = "Glucose")

# boxplots bmi by Glucose
ggplot(diabetes_v, aes(x = newBMI, y = Glucose)) +
  geom_boxplot(colour = "black", outline = FALSE) +
  theme + labs(title = "Graph 11. Distribution of Glucose by BMI", x = "BMI", y = "Glucose")

# boxplots bmi by SkinThickness 
ggplot(diabetes_v, aes(x = newBMI, y = SkinThickness)) +
  geom_boxplot(colour = "black", outlier = FALSE) +
  theme + labs(title = "Graph 12. Distribution of SkinThickness by BMI", x = "BMI", y = "SkinThickness") +
  ylim(0, 75)
```

```{r, fig.width=2}
# scatterplot age by glucose
ggplot(data = diabetes)+
  geom_point(mapping = aes(x = Age, y = Glucose), size = 0.5)+
  geom_smooth(mapping = aes(x = Age, y = Glucose), se=FALSE, color = "red")+
  theme + labs(title = "Graph 13. Regression of Glucose on Age", x = "Age", y = "Glucose")

# scatterplot bmi by glucose
ggplot(data = diabetes)+
  geom_point(mapping = aes(x = BMI, y = Glucose), size = 0.5)+
  geom_smooth(mapping = aes(x = BMI, y = Glucose), se=FALSE, color = "red")+
  theme + labs(title = "Graph 14. Regression of Glucose on BMI", x = "BMI", y = "Glucose")

# scatterplot DiabetesPedigreeFunction by glucose
ggplot(data = diabetes)+
  geom_point(mapping = aes(x = DiabetesPedigreeFunction, y = Glucose), size = 0.5)+
  geom_smooth(mapping = aes(x = DiabetesPedigreeFunction, y = Glucose), se=FALSE, color = "red")+
  theme + labs(title = "Graph 15. Regression of Glucose on DPF", x = "DiabetesPedigreeFunction", y = "Glucose")

```

```{r, fig.width=2}
# choose log(bmi) to predict glucose, justify that choice.
ggplot(data = diabetes)+
  geom_point(mapping = aes(x = log(BMI), y = Glucose), size = 0.5)+
  geom_smooth(mapping = aes(x = log(BMI), y = Glucose), se=FALSE, color = "red")+
  theme + labs(title = "Graph 16. Regression of Glucose on log(BMI)", x = "log(BMI)", y = "Glucose")
```


```{r, fig.width=2}
# choose log(age) to predict glucose, justify that choice.
ggplot(data = diabetes)+
  geom_point(mapping = aes(x = log(Age), y = Glucose), size = 0.5)+
  geom_smooth(mapping = aes(x = log(Age), y = Glucose), se=FALSE, color = "red")+
  theme + labs(title = "Graph 17. Regression of Glucose on log(Age)", x = "log(Age)", y = "Glucose")
```

##### Regression

```{r}
################################ linear regression part #############################
set.seed(1)
randOrder2 = order(runif(nrow(diabetes)))
train.df2 = subset(diabetes,randOrder2 < .8 * nrow(diabetes))
test.df2 = subset(diabetes,randOrder2 > .8 * nrow(diabetes))
# delete outcome in the datasettrain.df2
train.df2 <- train.df2[!names(train.df2) %in% c("Outcome")]
```

### correlation matrix

```{r, fig.width=6}
#plot the correlation matrix visual
cor2 <- cor(train.df2)
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(cor2, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )
```

```{r}
### forward
# create model with no predictors for bottom of search range
glu.lm.null <- lm(Glucose ~1, data = train.df2)
glu.lm <- lm(Glucose ~., data = train.df2)
# use step() to run forward selection
glu.lm.step_for <- step(glu.lm.null,   
                    scope=list(lower=glu.lm.null, upper=glu.lm), direction =  
                      "forward")
sum_for2 <- summary(glu.lm.step_for) 
sum_for2 

######
Glucose ~ newInsu + BloodPressure +  Age + DiabetesPedigreeFunction + BMI
```

```{r}
# backward
glu.lm.step_back <- step(glu.lm, direction = "backward")
sum_back2 <- summary(glu.lm.step_back) 
summary(glu.lm.step_back) 

#####
Glucose ~ newInsu + BloodPressure +  Age + DiabetesPedigreeFunction + BMI
```

```{r}
# both
glu.lm.step_both <- step(glu.lm, direction = "both")
sum_both2 <- summary(glu.lm.step_both) 
sum_both2

#####
Glucose ~ newInsu + BloodPressure +  Age + DiabetesPedigreeFunction + BMI
```

```{r}
# exhaustive
search.exhaustive2 <- regsubsets(Glucose ~ . , data = train.df2, nbest = 1, nvmax = dim(train.df2)[2],
                     method = "exhaustive")
sum_exhaustive2 <- summary(search.exhaustive2)
sum_exhaustive2

# show models
sum_exhaustive2$which
# show metrics
sum_exhaustive2$rsq;
sum_exhaustive2$adjr2
sum_exhaustive2$Cp

#####
Glucose ~ newInsu + BloodPressure +  Age + DiabetesPedigreeFunction + BMI
```

#### the best model for predict glucose
Glucose ~ DiabetesPedigreeFunction + BloodPressur + newInsu
```{r}
#  use options() to ensure numbers are not displayed in scientific notation.
options(scipen = 999)
Glucose_model<-lm(Glucose~BloodPressure + newInsu + BMI + DiabetesPedigreeFunction + Age,data=train.df2)
summary(Glucose_model)  # r^2 = 0.24
RMSE <- round(sqrt(c(crossprod(Glucose_model$residuals)) / length(Glucose_model$residuals)),2)
RMSE
#check for Variance Inflation Factor (VIF); must be < 10; should be less than 5
vif(Glucose_model)

## additional diagnostics to checsk for outliers/leverage points
par(mfrow=c(2,2))
plot(Glucose_model)

# remove outliers
train.df3 <- train.df2[-c(206,146,45,34,49),]

# use accuracy() to compute common accuracy measures. # rmse 26
accuracy(predict(Glucose_model,train.df3), train.df3$Glucose) %>% kable()
```

#### Validation
```{r}
#### Table 6.4
# use predict() to make predictions on a new set. 
glu.lm.pred <- predict(glu.lm, test.df2)
options(scipen=999, digits = 0)
residuals <- test.df2$Glucose - glu.lm.pred
result_glu<-data.frame("Predicted" = glu.lm.pred, "Actual" = test.df2$Glucose,
           "Residual" = residuals)

options(scipen=999, digits = 3)
# use accuracy() to compute common accuracy measures.
accuracy(glu.lm.pred, test.df2$Glucose) %>% kable()
```

```{r, fig.width=2}
## histogram for residuals
a<-data.frame(Glucose_model$residuals)
# Histogram with density plot
# Add mean line
p2<-ggplot(a, aes(x=Glucose_model.residuals)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666")+
  geom_vline(aes(xintercept=mean(Glucose_model.residuals)),
            color="red", linetype="dashed", size=1) +
  theme
p2
```

##### High-order
```{r}
# add high-order variable to regression
train.df2$exppedigree <- exp(train.df2$DiabetesPedigreeFunction)
train.df2$logbmi <- log(train.df2$BMI)
train.df2$logage <- log(train.df2$Age)

# new model with high-order variables
#options(scipen = 999)
Glucose_modelhi<-lm(Glucose~BloodPressure + logage +  logbmi + exppedigree + newInsu, data=train.df2)
sum_hi <- summary(Glucose_modelhi)

# exp(DiabetesPedigreeFunction) did not change a lot (so we keep this model as our best one)
Glucose_modelhi2<-lm(Glucose~BloodPressure + logage +  logbmi + DiabetesPedigreeFunction +
    newInsu,data=train.df2)
sum_hi2 <- summary(Glucose_modelhi2)

# show metrics
sum_hi  # adjust r^2 0.236, rmse 25.9
sum_hi2

RMSE <- round(sqrt(c(crossprod(sum_hi2$residuals)) / length(sum_hi2$residuals)),2)
RMSE
```

```{r}
# plot residuals and remove outliers (5/579 = 0.8%)
# plot residuaals for high-order model
plot(Glucose_modelhi2)
train.df3 <- train.df2[-c(248,175,58,503,43),]

Glucose_modelhi3<-lm(Glucose~BloodPressure + logage +  logbmi + DiabetesPedigreeFunction +
    newInsu,data=train.df3)
sum_hi3 <- summary(Glucose_modelhi3)
#### Validation of high-order model
# add high-order variable to regression
test.df2$exppedigree <- exp(test.df2$DiabetesPedigreeFunction)
test.df2$logbmi <- log(test.df2$BMI)
test.df2$logage <- log(test.df2$Age)

# use predict() to make predictions on a new set. 
glu.lm.pred2 <- predict(Glucose_modelhi2, test.df2)

# use accuracy() to compute common accuracy measures.
accuracy(glu.lm.pred2, test.df2$Glucose) %>% kable()
```

```{r}
################################ logistic regression part #############################
# CHANGE DATA TYPE
diabetes$Outcome <- as.factor(diabetes$Outcome)
diabetes$Pregnancies <- as.factor(diabetes$Pregnancies)

# divide data into train and test set
set.seed(1)
randOrder = order(runif(nrow(diabetes)))
train.df = subset(diabetes,randOrder < .8 * nrow(diabetes))
test.df = subset(diabetes,randOrder > .8 * nrow(diabetes))
```

##### correlation matrix

```{r, fig.width=6}
# plot the correlation matrix visual
train.df$Outcome <- as.numeric(train.df$Outcome)
train.df$Pregnancies <- as.numeric(train.df$Pregnancies)
cor <- cor(train.df)
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(cor, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )
```

```{r}
train.df$Outcome <- as.factor(train.df$Outcome)
train.df$Pregnancies <- as.factor(train.df$Pregnancies)
```

```{r}
### Forward Step-wise
# create model with no predictors for bottom of search range
dia.lm.null <- glm(Outcome~1, data = train.df, family = binomial)
dia.lm <- glm(Outcome~., data = train.df, family = binomial)
# use step() to run forward selection
dia.lm.step_for <- step(dia.lm.null,   
                    scope=list(lower=dia.lm.null, upper=dia.lm), direction = "forward")
sum_for <- summary(dia.lm.step_for) 

# Backward Step-wise
dia.lm.step_back <- step(dia.lm, direction = "backward")
sum_back <- summary(dia.lm.step_back) 

# Both Direction Step-wise
dia.lm.step_both <- step(dia.lm, direction = "both")
sum_both <- summary(dia.lm.step_both) 

# search
search <- regsubsets(Outcome ~ ., data = train.df, nbest = 1, 	nvmax = dim(train.df)[2], method = "exhaustive")
sum_sear <-summary(search)
sum_sear$which;
sum_sear$rsq;
sum_sear$adjr2;
sum_sear$Cp;

# comparison 
# same models with different methods
sum_for$coefficients
sum_back$coefficients
sum_both$coefficients

# best model with aic 536.4962
sum_for$aic

# Prediction on test data and accuracy test (73.1%)
tst_pred <- ifelse(predict(dia.lm.step_for, newdata = test.df, type = "response") > 0.5, "Yes", "No")
tst_tab <- table(predicted = tst_pred, actual = test.df$Outcome); sum(diag(tst_tab))/sum(tst_tab)
test_prob <- predict(dia.lm.step_for, newdata = test.df, type = "response")
test_roc <- roc(test.df$Outcome ~ test_prob, plot = TRUE, print.auc = TRUE) # 0.774
```

```{r, fig.width=3}
model1_data <- augment(dia.lm.step_for) %>% 
  mutate(index = 1:n()) %>%
  mutate(Outcome = ifelse(Outcome == "1", "0", "1"))
c <- ggplot(model1_data, aes(index, .std.resid, color = Outcome)) + 
  geom_point(stat = "identity") +
  labs(title = "Standardized Deviance Residuals", y = "Residual Std", x ="Residuals") +
  theme
c
```

